### 2022-10-28 src/Draft_Docs.Rmd
first commit, although plenty are done

---
title: "Draft"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up

```{r, message = FALSE}
library(e1071)
library(MASS)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tibble)
library(conflicted)
library(knitr)

# everytime `select` is called, it will prefer dplyr function instead of MASS
conflict_prefer(name = "select", winner = "dplyr")
conflict_prefer(name = "filter", winner = "dplyr")
```

## Data Cleaning
### MPI

```{r}
df.povert <- read.csv('../data/mpi.14.19.csv')
str(df.povert)
```
How many data point we have each year?
```{r}
no.avail <- df.povert %>%
  select(country.code, economy, year) %>% 
  group_by(year) %>% 
  summarise(`data points` = n())

kable(no.avail)

hist(df.povert$year, breaks = 2010:2022)
```
We want to find difference of MPI in 2 separate year (i.e. y1 and y2) in each country. It's useful to know how many countries which were investigated in y1, were also investigated in y2.
```{r}
# Extract data from a specific year
getYear <- function(yr) {
    df.povert %>% 
      select(country.code, economy, year, dep.monetary, dep.mpi) %>% 
      filter(year == yr)
}

# segmented data by year (2012-2021)
seg <- lapply(2012:2021, getYear)
names(seg) <- 2012:2021
str(seg)
```
Find overlapping countries
```{r}
# Calulate number of countries overlapping between s1 and s2
noOverlap <- function(y1, y2) {
  y1 %>%
    inner_join(y2, by = "country.code") %>% 
    filter(year.x != year.y) %>% 
    summarise(count = n()) %>% 
    pull()
}

# A 10x10 matrix to represent the no. of overlapping countries
cross <- diag(10)
rownames(cross) <- 2012:2021
colnames(cross) <- 2012:2021

# find no.of overlapping countries for each pair of years
for (i in 1:10) {
  cross[i,] = sapply(seg[1:10], noOverlap, y2 = seg[[i]])
}

cross
```
We find that each pair of years from 2012-2019 are fairly overlapping (mostly > 45). Therefore, taking any two years in this range will make a reasonable dataset for analysis.

Next, we make a function to return the difference in MPI in any two year
```{r}
mpiDiff <- function(y1, y2) {
  y1 <- as.character(y1)
  y2 <- as.character(y2)

  fin.df <- seg[[y1]] %>%
    inner_join(seg[[y2]], by = "country.code") %>% 
    filter(year.x != year.y) %>% 
    select(country.code,
           economy = economy.x,
           mpi.y1 = dep.mpi.x,
           mpi.y2 = dep.mpi.y) %>% 
    mutate(mpi.diff = mpi.y1 - mpi.y2)

  fin.df
}

# E.g. difference of mpi in 2012 and 2018
mpi.12.18 <- mpiDiff(2012, 2018)
head(mpi.12.18)
```
### Education
This is a very rough processing and not very rigorous. For reference only

First, do some summary on gdp (2012-2014)
```{r}
gdp <- read.csv('../data/gdp.csv')

gdp <- as.tibble(gdp) %>% 
  dplyr::select(economy, country.code, gdp.2012 = X2012, gdp.2013 = X2013, gdp.2014 = X2014)

gdp <- gdp %>% mutate(no.avail = 3 - rowSums(is.na(.)), mean.gdp = (gdp.2012 %+% gdp.2013 %+% gdp.2014) / no.avail) %>% 
  filter(!is.nan(mean.gdp)) %>% 
  dplyr::select(economy, country.code, mean.gdp)
```

```{r}
#custom plus operator to ignore NAs
#e.g. 1 + NA = 1.
`%+%` <- function(x, y) {
  mapply(sum, x, y, MoreArgs = list(na.rm = TRUE))
}

#import + clean data
edu.expense <- read.csv('../data/edu.percent.gdp.csv')
edu.expense <- as_tibble(edu.expense) %>% 
  dplyr::select(economy, country.code, edu2012 = X2012, edu2013 = X2013, edu2014 = X2014)

#get mean %expenditure on education 2012-2014
edu.expense <- edu.expense %>% mutate(no.avail = 3 - rowSums(is.na(.)), mean.edu = (edu2012 %+% edu2013 %+% edu2014) / no.avail) %>% 
  filter(!is.nan(mean.edu)) %>% 
  dplyr::select(economy, country.code, mean.edu)

# MPI with edu col
dat <- mpi.12.18 %>%
  dplyr::select(country.code, mpi.diff) %>% 
  inner_join(edu.expense %>% 
               dplyr::select(country.code, mean.edu), by = 'country.code') %>% 
  inner_join(gdp %>% 
               dplyr::select(country.code, mean.gdp), by = 'country.code')
head(dat)

lrm <- lm(mpi.diff ~ log(I(mean.edu * mean.gdp)), data = dat)
summary(lrm)
plot(dat$mean.gdp * dat$mean.edu, dat$mpi.diff)
plot(lrm, which=1)
hist(resid(lrm))
```