### 2022-10-28 src/outline.Rmd

---
title: "Miscellanous"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Structure
1. We find the impact of gov expenditure effect on poverty using conventional definition of poverty
2. We redo the analysis using MPI as indicator of poverty
3. We conclude form the results obtained.
4. Some surprising correlation we experimented with, and how these can be a product of different reasons
  1. Gender equality and poverty
  2.

## Reference

## Data source
[MPM](https://www.worldbank.org/en/topic/poverty/brief/multidimensional-poverty-measure)
[mpi](https://data.worldbank.org/indicator/SI.POV.MDIM)
[pov215](https://data.worldbank.org/indicator/SI.POV.DDAY)

[Country.Class](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)
[GDP.PC](https://data.worldbank.org/indicator/NY.GDP.PCAP.PP.CD)
[Pri.edu](https://data.worldbank.org/indicator/SE.XPD.PRIM.ZS)
[Sec.edu](https://data.worldbank.org/indicator/SE.XPD.SECO.ZS)
[ter.edu](https://data.worldbank.org/indicator/SE.XPD.TERT.ZS)
[total.edu](https://data.worldbank.org/indicator/SE.XPD.TOTL.GD.ZS)
[health.exp](https://data.worldbank.org/indicator/SH.XPD.GHED.GD.ZS)
[milis.exp](https://data.worldbank.org/indicator/MS.MIL.XPND.GD.ZS)

[fdi](https://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD)
[unemp](https://data.worldbank.org/indicator/SL.UEM.TOTL.NE.ZS)
[labor.frc.part](https://data.worldbank.org/indicator/SL.TLF.CACT.NE.ZS)
[gender.equal](https://data.worldbank.org/indicator/IQ.CPA.GNDR.XQ)
[pop.growth](https://data.worldbank.org/indicator/SP.POP.GROW)
[urb.grw](https://data.worldbank.org/indicator/SP.URB.GROW)
[rural.grw](https://data.worldbank.org/indicator/SP.RUR.TOTL.ZG)
[inflation](https://data.worldbank.org/indicator/NY.GDP.DEFL.KD.ZG)
[gross capital formation](https://data.worldbank.org/indicator/NE.GDI.TOTL.ZS)
[trade](https://data.worldbank.org/indicator/NE.TRD.GNFS.ZS)
[reg.class](https://datatopics.worldbank.org/world-development-indicators/the-world-by-income-and-region.html)
[income.class](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)