### 2022-11-03 src/All variables.Rmd
I created the linear regression model with all the variables

---
title: "Imtiyaaz"
output: html_document
date: "2022-11-03"
---
```{r}
library(tidyverse) # for data manipulations and visualisations
library(ggpubr) # for correlation coefficient on plot
library(gridExtra) # for arranging plots 
library(e1071) # for skewness
```

```{r}
df <-read.csv('data/countries.csv' , ,sep = "", header = TRUE)
df
subset1 <- df %>% select("income","edu.total","edu.pri","edu.sec","edu.ter","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","pop.gwth.urban","gdp.dflt","gdr.eql","gcf","trade")

```

```{r}
set.seed(10)
index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
train <- df[index, ]
test <- df[-index, ]
```

```{r}
lm.slr <- lm(pov ~ income+edu.total+edu.pri+edu.sec+edu.ter+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gdr.eql+gcf+trade,data = train)
summary(lm.slr)

```