### 2022-11-07 src/Selectively_remove_variables.Rmd
Create Selectively_remove_variables.Rmd
Updated my part on "Selectively removin g variables", tried to do ridge regression but realise that heteroskedascity still exists, will try to consult prof tmr

---
title: "Selectively Remove Variables"
author: "Callista Stephine Yu"
date: "2022-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
library(knitr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stats)
library(tidyverse)
library(e1071)
library(car)
```

```{r}
countries <- read.csv("../data/countries.csv")
```

```{r}
head(countries)
```

```{r}

#dataset without the red 
#removing reg as it is a bit too complex to deal for now, not relevant

#subset1 <- subset(countries, select = c("edu.total", "fdi", "gcf", "gdp.dflt", "hlth", "income", "lbr.part", "mil", "pop.gwth rural", "pop.gwth.urban", "pov", "reg", "trade", "unemp", "year"))
subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))

subset1 <- subset0 %>% mutate(lfdi = log(subset1$fdi)) %>% mutate(lgdp.pc = log(subset1$gdp.pc)) %>% mutate(lgdp.gdlt = log(subset1$gdp.dflt))
#dataset
```

```{r}
#converting non-numeric arguments into factors

subset1$income <- factor(subset1$income)
subset1$income <- as.numeric(subset1$income)

```

```{r}
#Perform standardisation

(df.countries0 <- subset1)
scaler <- apply(X = df.countries0, MARGIN = 2, FUN = sd)

(df.countries <- as.data.frame(apply(df.countries0, 2, function(x) x/sd(x, na.rm = TRUE))))

```

```{r}
#Divide each feature/target by its standard deviation

head(df.countries)
```

```{r}
#check for correlation

corrplot(cor(df.countries, use = "complete.obs"), 
         method = "number",
         type = "upper",
         tl.col = "black", 
         tl.srt = 30
         )
```

```{r}
set.seed(123)
index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
train_1 <- subset1[index,]
test_1 <- subset1[-index,]

```

```{r}
#Boxplot of the predictor variables
```

```{r}
```

```{r}
#Everything
lm_1 <- lm(data= train_1, formula = pov ~ .)
summary(lm_1)
```
```{r}
#remove hlth 
lm_2 <- lm(data= train_1, formula = pov ~ edu.total + gcf + fdi + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_2)
plot(lm_2, which = 5)
```

```{r}
#Remove fdi
lm_3 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_3)
plot(lm_3, which = 5)
```
```{r}
#remove year 
lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_4)
plot(lm_4, which = 5)
```
```{r}
#remove gdp.dflt 
lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_4)
plot(lm_4, which = 5)
```
```{r}
lm_5 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc)
summary(lm_5)
plot(lm_5, which = 5)
```
```{r}
#removing lfdi
lm_6 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc  + lgdp.pc)
summary(lm_6)
plot(lm_6, which = 4)
```
```{r}
#
data0 <- subset(df.countries, select = c("edu.total","gcf","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp","gdp.pc", 
                                        "lgdp.pc"))
data <- na.omit(data0)
train.x <- data.matrix(subset(data, select = -pov))
train.y = data.matrix(data$pov)

corrplot(cor(data, use = "complete.obs"), 
         method = "number",
         type = "upper",
         tl.col = "black", 
         tl.srt = 30
         )
```

```{r}
#Find the ridge regression model
set.seed(123)
rate <- 0.8
train.size <- round(nrow(data)*rate)
sample(sample(nrow(data), train.size))
training <- data[sample, ]
test <- data[-sample, ]

```

```{r}
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)

t(coef(glm_ridge))
```

```{r}
train.x <- model.matrix(pov ~., data = training)[, -1]
train.y <- training$pov
test.x <- model.matrix(pov ~., data = test)[,-1]
test.y <- test$pov
```

```{r}
model_ridge_trial1 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.1)
t(coef(model_ridge_trial1))
```

```{r}
model_ridge_trial2 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.5)
t(coef(model_ridge_trial2))
```

```{r}
model_ridge_trial3 <- glmnet(train.x, train.y, alpha = 0, lambda = 1)
t(coef(model_ridge_trial3))
```
```{r}
table_summary <- cbind(coef(model_ridge_trial1),
                       coef(model_ridge_trial2),
                       coef(model_ridge_trial3))

colnames(table_summary) <- c("lambda = 0.1",
                             "lambda = 0.5",
                             "lambda = 1")

table_summary
```

```{r}
#generate sequence of lambda values 
lambda <- 10^seq(-6, 2, length = 100)
ridge_model = glmnet(train.x, train.y, alpha = 0, lambda = lambda)

t(coef(ridge_model))
```

```{r}
plot(ridge_model, xvar = "lambda", label = TRUE)

#customise the plot with labels
add_lbs <- function(fit, offset_x = 2.5) {
  L <- length(fit$lambda[L]) + offset_x
  y <- fit$beta[,L]
  labs <- names(y)
  text(x, y , labels = labs, cex = 1.5)

  title(main = "Coefficients of Predictors vs Log(Lambda)",cex.main = 1.4, adj = 0, line = 3)
  title(sub = "Number of Predictors", cex.sub = 1.1, adj = 0.55, line = -21.5)
  add_lbs(ridge_model)
  legend("topright", lwd = 1, col = 1:6, legend = colnames(train.x), cex = .7)

}
```

```{r}
plot_glmnet(ridge_model)
```
```{r}
#cross - validation
set.seed(123)
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0, type.measure = "mse")

```

```{r}
cv_ridge
cv_ridge$lambda.min
cv_ridge$lambda.1se
```

```{r}
plot(cv_ridge)
abline(v=log(cv_ridge$lambda.min), col = "red", lty=5)
text(log(cv_ridge$lambda.min)+1, 1.4, 
     labels= paste0("lambda_min = ", 
                    round(cv_ridge$lambda.min, digits = 3)), 
     cex = 1.1)

# Label for lambda.1se
abline(v=log(cv_ridge$lambda.1se), 
       col = "red", lty=5)
text(log(cv_ridge$lambda.1se)+1, 1.25, 
     labels= paste0("lambda_1se = ", 
                    round(cv_ridge$lambda.1se, digits = 3)), cex = 1.1)

```

```{r}
#train the ridge regression model using lambda.min

glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
t(coef(glm_Ridge))
```
```{r}
eval_results <- function(fit, true) {
  actual <- data.matrix(true)
  SSE <- sum((actual - fit)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE/SST
  data.frame(
    MSE = MSE(fit, true),
    MAE = MAE(fit, true),
    RMSE = RMSE(fit, true),
    MAPE = MAPE(fit, true),
    R2  = R_square

  )
}
```

```{r}
fit <- predict(glm_Ridge, train.x)
true <- train.y
summary_Ridge_train <- eval_results(fit, true)
summary_Ridge_train
```

```{r}
fit <- predict(glm_Ridge, test.x)
true <- test.y
summary_Ridge_test <- eval_results(fit, true)
summary_Ridge_test
```

```{r}
summary_Ridge <- rbind(summary_Ridge_train[-5], summary_Ridge_test[-5])
rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
knitr::kable(summary_Ridge, digits = 3)

```

```{r}
#Obtain fitted and true values

fit <- predict(glm_Ridge, train.x)
true <- train.y

```

```{r}
#Plot residuals s
plot(fit, true - fit, 
     main = "Residual Plot of Ridge Regression Model on the Training dataset",
     xlab = "Fitted value",
     ylab = "Standardised Residuals")
abline(h = 0, col = "red", lwd = 2)
abline(h = 1, col = "red", lty = 2, lwd = 2)
abline(h = -1, col = "red", lty = 2, lwd = 2)

```
Constant varince assumption not satisfied. What to do? Most likely need to transform the variable

```{r}
mlr_adj <- lm(pov ~., data = training)

summary(mlr_adj)
```