### 2022-11-14 data/forward.Rmd

---
title: "Untitled"
output: html_document
date: "2022-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

```{r cars}
summary(lm(formula = pov ~ hlth,data = countries1))
summary(lm(formula = pov ~ fdi,data = countries1))
summary(lm(formula = pov ~ reg,data = countries1))
summary(lm(formula = pov ~ mil,data = countries1))
summary(lm(formula = pov ~ lbr.part,data = countries1))
summary(lm(formula = pov ~ unemp,data = countries1))
summary(lm(formula = pov ~ trade,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural,data = countries1))
summary(lm(formula = pov ~ pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ gcf,data = countries1))
summary(lm(formula = pov ~ gdp.dflt,data = countries1))
summary(lm(formula = pov ~ trade,data = countries1))
```
```{r}
summary(lm(formula = pov ~ hlth+reg,data = countries1))
summary(lm(formula = pov ~ fdi+reg,data = countries1))
summary(lm(formula = pov ~ mil+reg,data = countries1))
summary(lm(formula = pov ~ lbr.part+reg,data = countries1))
summary(lm(formula = pov ~ unemp+reg,data = countries1))
summary(lm(formula = pov ~ trade+reg,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total+reg,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural+reg,data = countries1))
summary(lm(formula = pov ~ pop.gwth.urban+reg,data = countries1))
summary(lm(formula = pov ~ gcf+reg,data = countries1))
summary(lm(formula = pov ~ gdp.dflt+reg,data = countries1))
summary(lm(formula = pov ~ trade+reg,data = countries1))
summary(lm(formula = pov ~ income+reg,data = countries1))

```
```{r}
summary(lm(formula = pov ~ income+reg,data = countries1))
summary(lm(formula = pov ~ hlth+reg+income,data = countries1))
summary(lm(formula = pov ~ fdi+reg+income,data = countries1))
summary(lm(formula = pov ~ mil+reg+income,data = countries1))
summary(lm(formula = pov ~ lbr.part+reg+income,data = countries1))
summary(lm(formula = pov ~ unemp+reg+income,data = countries1))
summary(lm(formula = pov ~ trade+reg+income,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total+reg+income,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural+reg+income,data = countries1))
summary(lm(formula = pov ~ pop.gwth.urban+reg+income,data = countries1))
summary(lm(formula = pov ~ gcf+reg+income,data = countries1))
summary(lm(formula = pov ~ gdp.dflt+reg+income,data = countries1))
summary(lm(formula = pov ~ trade+reg+income,data = countries1))
```
```{r}
summary(lm(formula = pov ~ mil + reg + income, data = countries1))

```
```{r}
summary(lm(formula = pov ~ pop.gwth.urban + reg + income + mil, data = countries1))
summary(lm(formula = pov ~ hlth+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ fdi+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ lbr.part+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ unemp+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ gcf+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ gdp.dflt+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
```
```{r}
countries9 <- countries1 %>% 
  select(Age, TotalWorkingYears, NumCourses, YearsAtCurrCompany, MonthlyIncome)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.