### 2022-11-16 data/forward.Rmd

---
title: "Untitled"
output: html_document
date: "2022-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df_count<-read.csv("../data/countries1.csv")
head(df_count)

set.seed(10)
index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
train <- df_count[index, ]
test <- df_count[-index, ]
```

Through forward selection, we will do feature selection. We will stop the forward selection when the adjusted r-value>0.75.

Through iteration of different models, a linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201.

Below are some of the linear models that have been explored.
```{r cars}
summary(lm(formula = pov ~ income,data = train)) #0.6201
summary(lm(formula = pov ~ reg,data = train)) #0.5569
summary(lm(formula = pov ~ edu.total,data = train)) #0.09577
summary(lm(formula = pov ~ hlth,data = train)) #0.0921
```
Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value. reg as a predictor has the highest adjusted r-value=0.7181

```{r}
summary(lm(formula = pov ~ reg+income,data = train)) #0.7181
summary(lm(formula = pov ~ edu.total+income,data = train)) #0.6033
summary(lm(formula = pov ~ hlth+income,data = train)) #0.6201
summary(lm(formula = pov ~ mil+income,data = train)) #0.6282
```

The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7492
```{r}
summary(lm(formula = pov ~ edu.total+income+reg,data = train)) #0.7347
summary(lm(formula = pov ~ hlth+income+reg,data = train)) #0.6201
summary(lm(formula = pov ~ mil+income+reg,data = train)) #0.6282
summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train)) #0.7492
```
The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
```{r}
summary(lm(formula = pov ~ edu.total+income+reg+lgdp.pc,data = train)) #0.7495
summary(lm(formula = pov ~ mil+income+reg+lgdp.pc,data = train)) #0.7571
summary(lm(formula = pov ~ pop.gwth.urban+income+reg+lgdp.pc,data = train)) #0.7606
summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7703
```

Fitting of multiple linear regression with predictor gdp.pc, income, reg and lgd.pc.
```{r}
lm1<-lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)
summary(lm1)
```

Assessment of linear model lm1: 
the predicted model has some NA values which are removed. The error rates are too high, therefore, this method of variable selection may not be suitable
```{r}
eval.metrics.linreg <- function(actual, predicted) {
  residual <- actual - predicted
  mse <- mean(residual ^ 2)
  mae <- mean(abs(residual))
  rmse <-  sqrt(mse)
  mape <- mean(abs(residual / actual)) * 100

  data.frame(
    MSE = mse,
    MAE = mae,
    RMSE = rmse,
    MAPE = mape
  )
}
predicted<-predict(lm1,newdata=test)
actual<-test$pov
df23<-na.omit(data.frame(actual,predicted))
ac<-df23$actual
pr<-df23$predicted
eval.metrics.linreg(ac,pr)
```

Intepretation: gdp per capital, income class and region of the country can provide a model to predict the poverty headcount ratio in a country.