### 2022-11-16 src/All variables.Rmd
Update All variables.Rmd

---
title: "Imtiyaaz"
output: html_document
date: "2022-11-03"
---
```{r}
library(tidyverse) # for data manipulations and visualisations
library(ggpubr) # for correlation coefficient on plot
library(gridExtra) # for arranging plots 
library(e1071) # for skewness
library(ggplot2)
library(dplyr)
library(tidyr)
library(readxl)
library(corrplot)
library(car)
library(DescTools)
library(glmnet)
```

First we will write the function "eval_results(true,predict) to evaluate the models
```{r}
eval_results <- function(true, predict) {
  actual <- data.matrix(true)
  SSE <- sum((predict - actual)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE / SST
  data.frame(
    MSE = MSE(predict, true),
    MAE = MAE(predict, true),
    RMSE = RMSE(predict, true),
    Rsquare = R_square
  )
}
```
First we read the csv file for the countries containing the dataset. We will then put the variables we need in to subset1.
```{r}
df <-read.csv('../data/countries1.csv')
df
subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
subset1
```

Next we will build a SLR model. We will put 80% of dataset to train and the rest to test.
```{r}
set.seed(10)
index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
train <- df[index, ]
test <- df[-index, ]
```

The model output will be placed in lm.slr and can be viewed in summary(lm.slr). THe r-squared value is 0.9619.
```{r}
lm.slr <- lm(pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
summary(lm.slr)

```
Next we will be perfoming regularisation. We have to use this function as not all the varibales are numerical. Therefore to overcome it we build the function below.
```{r}
subset1$income <- factor(subset1$income)
subset1$country.code <- factor(subset1$country.code)

sd0 <- function(vct) {
  if (!is.numeric(vct)) {
    return(NA)
  }

  return(sd(vct, na.rm = T))
}

std0 <- function(vct, scl) {
  if (!is.numeric(vct)) {
    return(vct)
  }

  return(vct/scl)
}

scale <- function(data) {
  unlist(lapply(data, sd0))
}

standardise <- function(data) {
  scaler <- scale(data)

  numCols <- which(unlist(lapply(data, is.numeric)))
  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
  fct <- data[,-numCols]
  return(cbind(fct, num))
}
```

Next we will perform standardistation to the dataset.
```{r}
str(subset1)
df0 <- subset1
subset1
apply(subset1, 2, sd)

```

Next we will plot a pov vs income scatter plot to see the effect of income on pov. From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov. This is consistent with our hypothesis that countries that are richer do not have as much poverty issue as the poorer countries.
```{r}
ggplot(data = subset1, aes(x = income, y = pov)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Pov vs. Income ",
       x = "income",
       y = "pov")

```

```{r}
model1 <- lm(pov ~ ., data = subset1)
summary(model1)
vif(model1)

```
```{r}
subset2 <- subset(subset1,select =c("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
subset2 <- na.omit(subset2)
subset2
train.x <- data.matrix(subset(subset2, select = -pov))
train.y = data.matrix(subset2$pov)
```

```{r}
split_train_test <- function(data) {
set.seed(1984)
# split data
seeds <- data %>% 
  group_by(country.code) %>% 
  filter(row_number() == 1)

plants <- data %>% 
  group_by(country.code) %>% 
  filter(row_number() != 1)

isComplete <- which(complete.cases(plants))
idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))

training <- plants[-idx,] %>% 
  rbind(seeds)
test <- plants[idx,]

return(list(training, test))
}
```

```{r}

training <- split_train_test(subset2)[[1]]
test <- split_train_test(subset2)[[2]]
```

```{r}
ntrain <- 1:nrow(training)
combine <- rbind(training, test)
combine.mtrx <- model.matrix(pov ~ ., data = combine)
train.mtrx <- combine.mtrx[ntrain,]
test.mtrx <- combine.mtrx[-ntrain,]

train.x <- train.mtrx[,-1]
train.y <- training$pov

test.x <- test.mtrx[,-1]
text.y <- test$pov
```

```{r}
model3 <- lm(pov ~., data = training)
summary(model3)

```

```{r}
fit <- model3$fitted.values
true <- training$pov
summary_MLR3_train <- eval_results(true, fit)
```

```{r}
fit <- predict(model3, newdata=test)
true <- test$pov
summary_MLR3_test <- eval_results(true, fit)
```

```{r}
summary_MLR <- rbind(summary_MLR3_train[-5], 
                     summary_MLR3_test[-5])
rownames(summary_MLR) <- c("Baseline MLR_Train", 
                           "Baseline MLR_Test" )
knitr::kable(summary_MLR, digits = 3)
```

```{r}
# Training Data
data <- training
train.x <- data.matrix(data[,-ncol(data)])
train.y <- data.matrix(data[,ncol(data)])

# Test Data
data <- test
test.x <- data.matrix(data[,-ncol(data)])
test.y <- data.matrix(data[,ncol(data)])
```

```{r}
# Fit Ridge Regression Model
set.seed(123)
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
glm_ridge <- glmnet(train.x, train.y, 
                    alpha = 0, lambda = cv_ridge$lambda.min)

# Evaluate Ridge Regression Model on Training Set
fit <- predict(glm_ridge, newx = train.x, type = "response")
true <- train.y
summary_ridge_train <- eval_results(true, fit)

# Evaluate Ridge Regression Model on Test Set
fit <- predict(glm_ridge, newx = test.x, type = "response")
true <- test.y
summary_ridge_test <- eval_results(true, fit)

# Summarise Ridge Regression Model Performance into Table
summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
```

```{r}
# Fit LASSO Model
set.seed(123)
cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
                    lambda = cv_lasso$lambda.min)

# Evaluate LASSO Model on Training Set
fit <- predict(glm_lasso, newx = train.x)
true <- train.y
summary_lasso_train <- eval_results(true, fit)

# Evaluate LASSO Model on Test Set
fit <- predict(glm_lasso, newx = test.x)
true <- test.y
summary_lasso_test <- eval_results(true, fit)

# Summarise LASSO Model Performance into Table
summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
```

```{r}
summary_3models <- rbind(summary_MLR3_train, 
                         summary_ridge_train,                               
                         summary_lasso_train, 
                         summary_MLR3_test,                             
                         summary_ridge_test, 
                         summary_lasso_test)

rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
                               "LASSO Model_Train",  "Baseline MLR_Test",  
                               "Ridge Model_Test", "LASSO Model_Test" )

knitr::kable(summary_3models[1:3,-5], digits = 3)
knitr::kable(summary_3models[4:6,-5], digits = 3)

```